Non-word repetition in Yélî Dnye

### 2020-12-23 manuscript.tex
pass in progress, knits

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{NWE in Yélî Dnye}
\author{Alejandrina Cristia\textsuperscript{1}\ \& Marisa Casillas\textsuperscript{2,3}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Laboratoire de Sciences Cognitives et de Psycholinguistique, Département d'Etudes cognitives, ENS, EHESS, CNRS, PSL University\\\textsuperscript{2} Max Planck Institute for Psycholinguistics\\\textsuperscript{3} University of Chicago}
\authornote{All data are made available in a repository in the Open Science Framework. AC acknowledges the support of the Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004, ANR-14-CE30-0003, ANR-17-EURE-0017); and the J. S. McDonnell Foundation Understanding Human Cognition Scholar Award.

Correspondence concerning this article should be addressed to Alejandrina Cristia, 29, rue d’Ulm, 75005 Paris, France. E-mail: alecristia@gmail.com}
\keywords{phonology, non-word repetition, development\newline\indent Word count: ~xxx words}
\usepackage{lineno}

@@

\abstract{
In nonword repetition (NWR) studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in their language. NWR scores are thought to reflect long-term phonological knowledge as well as online phonological working memory and flexible production patterns. In this study, we report on NWR results among children learning Yélî Dnye, an isolate spoken on Rossel Island in Papua New Guinea that has an unusually dense phonological inventory. This study contributes to three lines of research. First, we document that non-word items containing typologically frequent sounds are repeated without changes more often that non-words containing typologically rare sounds. Second, we document rather weak effects of item length, contributing to other research suggesting that length effects may be language-specific. Third, we do not find strong individual variation effects in this population, contrary to previous results documenting strong age-related effects. Together, these data provide a unique view of online phonological processing in an understudied language while making preliminary connections between patterns in language development and in the frequency of linguistic features in human language at large.
}

@@

\hypertarget{todo-alex}{%
\subsection{TODO Alex}\label{todo-alex}}

\begin{itemize}
\tightlist
\item
  check whether it's stimuli, stimuli2, stimuli3 that should be used!!!
\item
  create fields final-data DONE
\item
  leave only files actually needed DONE
\item
  add readme explaining files DONE
\item
  remove all files that aren't necessary DONE
\item
  get rid of all the conversions and directly read in xls
\item
  set up OSF (or revise if already there)
\item
  implement decisions on prop phones \& patterns nwr mp
\item
  check and accept/deny MC's updated segment frequencies
\item
  re-run analyses w/ new frequencies and check that outcomes didn't qualitatively change
\item
  check YD data in figure 1 (the effect of item length looks reversed)
\item
  see MC comments throughout and delete now-irrelevant things
\end{itemize}

\hypertarget{todo-middy}{%
\subsection{TODO Middy}\label{todo-middy}}

@@

Children's perception and production of phonetic and phonological units continues developing well beyond the first year of life, even extending into middle childhood (e.g., Hazan \& Barrett, 2000). Much of the evidence for later phonological development comes through the use of nonword repetition (NWR) tasks. In a NWR task, participants hear a short word-like form that is phonologically legal but lexically meaningless in the language(s) they are learning. After hearing this non-word, the participant's task is to try to immediately and precisely repeat it. NWR scores are thought to reflect long-term phonological knowledge (to perceive the item precisely despite not having heard it before) as well as online phonological working memory (to encode the item in the interval between hearing it and saying it back) and flexible production patterns (to produce the item precisely despite not having pronounced it before).

NWR has been used to seek answers to a variety of theoretical questions, including what the links are between phonology, working memory, and the lexicon (Bowey, 2001), and how extensively phonological constraints found in the lexicon affect online production (Gallagher, 2014). NWR is also frequently used in applied contexts, notably as a diagnostic tool for language delays and disorders (Estes, Evans, \& Else-Quest, 2007). Since non-words can be generated in any language, it has attracted the attention of researchers working in multilingual and linguistically diverse environments, particularly in Europe (Action, 2009; Meir, Walters, \& Armon-Lotem, 2016). However, it has seldom been used outside of Europe and North America (with exceptions including Gallagher, 2014, and @cristia2020infant). Further, the languages it has been used with have historically had small to moderate phonological inventories (Maddieson, 2005).

In the present study, we use NWR to investigate the phonological development of children learning Yélî Dnye, an isolate language spoken in Papua New Guinea (PNG) that has a large and unusually dense phonological inventory. The study was designed to contribute to multiple aspects of our understanding of phonological development: First, we included a subset of non-word items with typologically rare and/or challenging sounds to ask whether these rare sounds are disadvantaged in perception and/or production, both in terms of overall repetition scores and patterns of mispronunciation. Second, we varied the length (in syllables) of non-words to investigate the impact of word length on repetition and mispronunciation; previous NWR research has shown variable effects of length between populations. Third, these data give a first impression of how this task is responded to in a rarely-studied language and cultural setting; we add to this contribution by further investigating whether there are structured sources that account for individual variation in our sample.

Before jumping into the details of our study design we first give an overview of Yélî Dnye phonology as well as a brief ethnographic review of the developmental environment on Rossel Island. Since NWR has been almost exclusively used in urban, industrialized populations (see Cristia et al., 2020 for an exception), we provide this additional ethographic information to contextualize the adaptations we have made in running the task and in gathering age and other demographic information, compared to what is typical in urban, industrialized settings.

@@

Yélî Dnye is an isolate language (presumed Papuan) spoken by approximately 7,000 people residing on Rossel Island, an island found at the far end of the Louisiade Archipelago in Milne Bay Province, Papua New Guinea. The Yélî sound system, much like its baroque grammatical system (Levinson, n.d.), is unlike any other in the region.

With only four primary places of articulation (bilabial, alveolar, post-alveolar, and velar) and no voicing contrasts, the phonological inventory is remarkably packed with acoustically similar segments. The core oral stop set includes both singleton (/p/, /t/, /ṭ/, and /k/) and doubly-articulated (/tp/, /ṭp/, /kp/) segments, with full nasal equivalents (/m/, /n/, /ṇ/, /ŋ/, /nm/, /ṇm/, /ŋm/), and with a substantial portion of them contrastively pre-nasalized or nasally released (/mp/, /nt/, /ṇṭ/, /ŋk/, /nmtp/, /ṇmṭp/, /ŋmkp/, /ṭṇ/, /kŋ/, /ṭpṇm/, /kpŋm/). A large number of this combinatorial set can further be contrastively labialized, palatalized on release, or both (e.g., /pʲ/, /pʷ/, /pʲʷ/; /tpʲ/; /ṇmḍbʲ/; see Levinson (n.d.) for details).\footnote{We use Levinson's (n.d.) under-dot notation (e.g., /ṭ/) to indicate the post-alveolar place of articulation; these stops are, articulatorily, somewhat variable in place, with at least some tokens produced fully sub-apically. In approximating cross-linguistic segment frequency below we use the corresponding retroflex for each stop segment (e.g., /ʈ/, /ʈp/, /ɳ/).} The consonantal inventory also includes a number of non-nasal continuants (/w/, /j/, /ɣ/, /l/, /βʲ/, /lʲ/, /lβʲ/). Vowels in Yélî Dnye may be oral or nasal, short or long. The 10 oral vowel qualities, which span four levels of vowel height, (/i/, /ɯ/, /u/, /e/, /o/, /ə/, /ɛ/, /ɔ/, /æ/, /ɑ/) can be produced as short and long vowels, with 7 of these able to appear as short and long nasal vowels as well /ĩ/, /ũ/, /ə̃/, /ɛ̃/, /ɔ̃/, /æ̃/, /ɑ̃/).

In total, Yélî Dnye uses 90 distinctive segments (not including an additional 3 rarely used consonants), far outstripping the phonemic inventory size of other documented Papuan languages (Foley, 1986; Levinson, n.d.) and includes at least two contrasts not yet been documented elsewhere (labial-coronal double-articulations with dental vs.~post-alveolar coronal placement for both oral and nasal stops, Ladefoged \& Maddieson, 1996; Maddieson \& Levinson, n.d.).

Most words in Yélî Dnye are bisyllabic (\textasciitilde{}50\%), with monosyllabic words (\textasciitilde{}40\%) appearing most commonly after that, and with tri-and-above syllabic words appearing least frequently (\textasciitilde{}10\%; based on \textgreater{}5800 lexemes in the most recent Yélî Dnye dictionary at time of writing; Levinson, n.d.). The vast majority of syllables use a CV format. A small portion of the lexicon features words with a final closed syllable, limited to codas of -/m/, -/p/, or -/j/ (e.g., \enquote{ndap} (Spondylus shell) /ṇṭæp/). However, in spontaneous speech an epenthetic /ɯ/ is often inserted after word-final coda, yielding a CV.CV structure instead (e.g., \enquote{ndapî} /\enquote{ṇṭæ.pɯ/). This process is used frequently with English loan words that have a coda (e.g., 'ponî} (phone) /\enquote{pɔ.nɯ/). An even smaller portion of the lexicon features words starting with a vowel (e.g., 'ala} (here) /æ.'læ/), but these are limited to /æ/-. Finally, the lexicon features a handful of single-vowel grammatical morphemes (see Levinson (n.d.) for details).

@@

In a basic NWR task, the participant listens to a production of a word-like form, such as /bilik/, and then repeats back what they heard without changing any phonological feature that is contrastive in the language. For instance, in English, a response of {[}bilig{]} or {[}pilik{]} would be scored as incorrect; a response {[}biːlik{]}, where the vowel is lengthened without change of quality would be scored as correct, because English does not have a general feature of contrastive vowel length. There is some variation in how past NWR studies have designed the presentation procedure and structure of items. For example, while items are often presented orally by the experimenter (Torrington Eaton, Newman, Ratner, \& Rowe, 2015), an increasing number of studies have turned instead to playing back pre-recorded stimuli in order to increase control in stimulus presentation (Brandeker \& Thordardottir, 2015). Additionally, while some studies have used 10-15 non-words, others have employed up to 46 unique items (Piazzalunga, Previtali, Pozzoli, Scarponi, \& Schindler, 2019). Authors also often modulate structural complexity, typically measured in terms of item length (measured in number of syllables) and/or syllable structure (open as opposed to closed syllables, Gallon, Harris, \& Van der Lely, 2007).

@@

In view of the hypothesis-driven nature of this work, we had considered boosting the interpretational value of this evidence by announcing our analysis plans prior to conducting them. However, we realized that even pre-registering an analysis would be equivocal because we would not have enough power to look at all relationships of interest, in many cases possibly not enough to detect any of the known effects, given their variability across studies. To illustrate this issue, we portray in Figure 1 studies in which children's NWR scores were gathered between 4 and 12 years of age, and reported separately for items that are relatively short (1-2 syllables) versus longer items (3-4 syllables). Notice that the effect of stimulus length is minuscule among Italian children (Piazzalunga et al., 2019), but considerable among Tsimane' children (Cristia et al., 2020), where a drop of 40 percentage points is observed at all ages. A similar large difference in NWR scores for short versus long items was observed among Arabic children (Jaber-Awida, 2018). Even the effect of age is unstable in this sample of studies. Whereas it is quite clear that children's NWR scores increase in the Italian data, age effects are less stable among Tsimane' children. Therefore, all analyses in the present study are descriptive and should be considered exploratory.

@@

A second list of candidate items was generated in 2019 by selecting complex consonants and systematically crossing them with all the vowels in the Yélî Dnye inventory to produce consonant-vowel monosyllabic forms. As before, items were automatically excluded if they appeared in the dictionary. Additionally, since perceiving vowel length in isolated monosyllables is challenging, any item that had a short/long lexical neighbor was excluded. Because there is still much to discover about the phonology and phonetics of Yélî Dnye (Levinson, n.d.), it was also possible that we initially generated items with illegal, but currently undocumented constraints. Therefore, we made sure that the precise consonant-vowel sequence occurred in some real word in the dictionary (i.e., that there was a longer word included the monosyllable as a subsequence). These candidates were then presented to one informant, for a final check that they did not mean anything. Together with the 2018 selection, they were recorded, based on their orthographic forms, using a Shure SM10A XLR dynamic headband microphone and an Olympus WS-832 stereo audio recorder (using an XLR to mini-jack adapter) by the same informant, monitored by the second author for clear production of the phonological target. The complete recorded list was finally presented to two more informants, who were able to repeat all the items and who confirmed there were no real words present. Despite these checks, one monosyllable was ultimately frequently identified as a real word in the resulting data (intended \enquote{yî} /yɯ/; identified as \enquote{yi} /yi/, \emph{tree}). This item is removed from the analyses below.

@@

We ran children in batches, testing within a handful of hamlet clusters spread across the northeastern region of the island. Because space availability was limited in different ways from hamlet to hamlet, the places where elicitation happened varied across testing sites. We tested in four different sites, only making a single visit to each, conducting back-to-back testing of all eligible children present at the time of our visit in order to prevent the items from \enquote{spreading} between children through hearsay. In the first hamlet, we tested children in five different places, with some children being tested inside a house and others tested on the veranda. The complete list of places and the ways in which they met the desiderata mentioned above can be found in the raw data, available from the online supplementary materials. Whenever children living in the same household were tested, we tried to test children in age order, from oldest to youngest, to minimize intimidation for younger household members, and always using different elicitation sets.

@@

Additionally, we estimated the typological frequency of all phonological segments used in the target items using the PHOIBLE cross-linguistic phonological inventory database (Moran \& McCloy, 2019). For each phone in our task, we extracted the number and percentage of languages noted to have that phone in its inventory. While PHOIBLE is an unprecedentedly comprehensive database, with phonological inventory data for over 2000 languages at time of writing, it is of course still far from complete, so estimated frequencies should be taken with a pinch of salt. In both PHOIBLE and UPSID---another large phonemic inventory database---note that nearly half of the segment types are only attested in one language (Steven Moran, personal communication; see also \href{http://web.phonetik.uni-frankfurt.de/upsid_info.html}{web.phonetik.uni-frankfurt.de/upsid\_info.html}). It is therefore unsurprising that we did not find three of our segments (/lβʲ/, /ṭp/, and /tp/) in the database. Following the general pattern that many phones only appear once cross-linguistically, even in large databases, we treat our \enquote{missing} phones as having a frequency of 1, with a (rounded) percentage of 0\%.

@@

We next checked whether the association between whole-item NWR scores and cross-linguistic phone frequence could actually be due to frequency of the sounds within the language: One can suppose that sounds that occur more frequently across languages are also more frequent within a language, and therefore may be easier for children to represent and repeat. We estimated frequency of the phones present in the stimuli in a corpus of child-centered recordings (Casillas et al., 2020) by counting the number of word types in which they occurred, and applied the natural logarithm.\footnote{We also carried out this analysis using token phone frequency, but this measure was not correlated with whole-item NWR scores, and therefore the fact that it did not explain away the predictive value of cross-linguistic phone frequency was less informative than the relationship discussed in the main text, with type frequencies.} As before, unattested sounds were not considered (i.e., they were declared NA so that they do not count for analyses). Phone frequencies estimated from this corpus were correlated with cross-linguistic phone frequencies (r(27)=0.50, p \textless{} 0.01). Moreover, averaging across items and participants, the Pearson correlation between scaled average corpus phone frequency and whole-item NWR scores was r = 0.46. Therefore, we fit another mixed logistic regression, this time declaring as fixed effects both scaled cross-linguistic and corpus frequencies (averaged across all attested phones within each stimulus item), in addition to age. As before, the model contained random slopes for both child ID and target. In this model, both cross-linguistic frequency (ß = 0.82, SE ß = 0.27, p \textless{} 0.01) and age (ß = 0.33, SE ß = 0.13, p = 0.01) were significant predictors of whole-item NWR scores, but corpus frequency (ß = 0.01, SE ß = 0.25, p = 0.98) was not.

@@

Finally, we looked into the frequency with which mispronunciations resulted in real words. In fact, two thirds of incorrect repetitions were recognizable as real words or phrases in Yélî Dnye or English: 63\%. This type of analysis is seldom reported---we could only find one comparison point: ({\textbf{???}}) found that illiterate Portuguese adults' NWR mispronunciations resulted in real words in 11.16\% of cases, whereas literate participants did so in only 1.71\% of cases. The percentage we observe here is much higher than reported in Castro and colleagues' study, but we do not know whether age, language, or even test structure explains this difference.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-is08042009language}{}%
Action, C. (2009). Language impairment in a multilingual society: Linguistic patterns and the road to assessment. \emph{Brussels: COST Office. Available Online at: Http://Www. Bi-Sli. Org}.

\leavevmode\hypertarget{ref-balladares2016socio}{}%
Balladares, J., Marshall, C., \& Griffiths, Y. (2016). Socio-economic status affects sentence repetition, but not non-word repetition, in Chilean preschoolers. \emph{First Language}, \emph{36}(3), 338--351. \url{https://doi.org/10.1177/0142723715626067}

@@

@@

>>>>

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{NWE in Yélî Dnye}
\author{Alejandrina Cristia\textsuperscript{1}\ \& Marisa Casillas\textsuperscript{2,3}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Laboratoire de Sciences Cognitives et de Psycholinguistique, Département d'Etudes cognitives, ENS, EHESS, CNRS, PSL University\\\textsuperscript{2} Max Planck Institute for Psycholinguistics\\\textsuperscript{3} University of Chicago}
\authornote{

Correspondence concerning this article should be addressed to Alejandrina Cristia, 29, rue d’Ulm, 75005 Paris, France. E-mail: alecristia@gmail.com}
\keywords{phonology, non-word repetition, development\newline\indent Word count: ~9,000 words}
\usepackage{lineno}

@@

\abstract{
In nonword repetition (NWR) studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in their language, and asked to repeat this item as closely as possible. NWR scores are thought to reflect some aspects of phonological development, saliently a perception-production loop supporting flexible production patterns. In this study, we report on NWR results among children learning Yélî Dnye, an isolate spoken on Rossel Island in Papua New Guinea. This study contributes to four lines of research. First, we document that non-word items containing typologically frequent sounds are repeated without changes more often that non-words containing typologically rare sounds, above and beyond any within-language frequency effects. Second, we document rather weak effects of item length, contributing to mounting research suggesting that length effects may be language- or population-specific. Third, we do not find strong individual variation effects in this population. Fourth, we discuss how NWR scores can be compared across studies, languages, and populations, in order to shed light on the factors universally structuring variation in language development. Together, these data provide a unique view of online phonological processing in an understudied language while making preliminary connections between language development and cross-linguistic features.
}

@@

@@

Children's perception and production of phonetic and phonological units continues developing well beyond the first year of life, even extending into middle childhood (e.g., Hazan \& Barrett, 2000). Much of the evidence for later phonological development comes from nonword repetition (NWR) tasks. In a NWR task, participants hear a short word-like form that is phonologically legal but lexically meaningless in the language(s) they are learning. After hearing this non-word, the participant's task is to try to immediately and precisely repeat it. NWR scores are thought to reflect long-term phonological knowledge (to perceive the item precisely despite not having heard it before) as well as online phonological working memory (to encode the item in the interval between hearing it and saying it back) and flexible production patterns (to produce the item precisely despite not having pronounced it before).

NWR has been used to seek answers to a variety of theoretical questions, including what the links are between phonology, working memory, and the lexicon (Bowey, 2001), and how extensively phonological constraints found in the lexicon affect online production (Gallagher, 2014). NWR is also frequently used in applied contexts, notably as a diagnostic tool for language delays and disorders (Estes, Evans, \& Else-Quest, 2007). Since non-words can be generated in any language, it has attracted the attention of researchers working in multilingual and linguistically diverse environments, particularly in Europe (COST Action, 2009; Meir, Walters, \& Armon-Lotem, 2016).

In the present study, we use NWR to investigate the phonological development of children learning Yélî Dnye, an isolate language spoken in Papua New Guinea (PNG) that has a large and unusually dense phonological inventory. The study was designed to contribute to four aspects of our understanding of phonological development.

First, we included a subset of non-word items with typologically rare and/or challenging sounds to ask whether these rare sounds are disadvantaged in the perception-production loop involved in NWR. Previous work using NWR has preferred relatively universal and early-acquired phonemes (with the possible exception of Gallagher, 2014), likely as a way to separate phoneme pronunciation from broader syllable structure and word-level prosodic effects (Gallon, Harris, \& Van der Lely, 2007). Here, we investigate repetition of non-word items containing cross-linguistically common and cross-linguistically rare phonetic targets.

Second, we varied the length (in syllables) of non-words to contribute to growing research looking at the impact of word length on NWR repetition, and what this may reflect about phonological development. Our reading of previous NWR research is that there are variable effects of length between populations. For instance, Jaber-Awida (2018) reports an average of 95\% correct repetition for items 1 or 2 syllables long among Arabic-speaking children at about 5 years of age, but 55\% for items 3 or 4 syllables long. In contrast, Piazzalunga, Previtali, Pozzoli, Scarponi, and Schindler (2019) observe nearly no change in performance in similarly-aged Italian learners, with a score of 84\% for 1-2 syllables versus 85\% for 3-4 syllables. It is possible that differences are due to a host of variables, including the modal length of words in the language and/or in child-directed speech in that culture. That said, there could be other causal pathways: Research on adults suggests that illiterate Portuguese speakers repeat monosyllabic non-words just as accurately as literate adults, whereas scores are much lower among illiterate than literate speakers for items 3 or 4 syllables long (de Santos Loureiro et al., 2004). Given that both groups of adults speak the same language (Brazilian Portuguese), then perhaps differences in repetition accuracy reveal differences in how flexible the perception-production loop is. Given the paucity of evidence looking at this question, we do not approach this issue within a hypothesis-testing framework but sought instead to provide one more piece of data on the question, which may be re-used in future meta- or mega-analytic approaches.

Third, there are ongoing discussions as to what the key factors structuring individual variation are. Although the ideal systematic review is missing, a recent paper comes close with a rather extensive review on the literature looking at correlations between NWR scores and a variety of child-level variables (Farabolini, Rinaldi, Caselli, \& Cristia, 2021). In a nutshell, most evidence is mixed, suggesting that consistent individual variation effects may be small, and more data is needed to estimate their true size. For this reason, we descriptively report association strength between NWR scores and child age, sex, birth order, and maternal education.

Fourth, these data contribute to the small literature using this task with non-Western, non-urban populations, speaking a language with a moderate to large phonological inventory (see Maddieson, 2005 for a broad classification of languages based on inventory size). Indeed, NWR has seldom been used outside of Europe and North America (with exceptions including Gallagher, 2014; Cristia, Farabolini, Scaff, Havron, \& Stieglitz, 2020), outside urban settings (except for in Cristia et al., 2020), nor with languages having large phonological inventories. There are no theoretical reasons to presume that the technique will not generalize to these new conditions. That said, Cristia et al. (2020) recently reported relatively lower NWR scores among the Tsimane', a non-Western rural population, interpreting these findings as consistent with the hypothesis that lower levels of infant-directed speech and/or low prevalence of literacy in a population could lead to population-level differences in NWR scores. In view of these results, it is important to bear in mind that NWR is a task developed in countries where literacy is widespread, and it is considered an excellent predictor of reading, for instance better than rhyme awareness (e.g., Gathercole, Willis, \& Baddeley, 1991). Therefore, it may not be a general index of phonological development, but more specifically reflect certain skills. Indeed, Cristia et al. (2020) present the task as being a good index of the development of \enquote{short-hand-like} representations specifically, which could thus miss, for example, more holistic representations.

Before going into the details of our study design we first give an overview of Yélî Dnye phonology as well as a brief ethnographic review of the developmental environment on Rossel Island. As discussed above, NWR has been almost exclusively used in urban, industrialized populations, so we provide this additional ethographic information to contextualize the adaptations we have made in running the task and in gathering age and other demographic information, compared to what is typical in urban, industrialized settings.

@@

Yélî Dnye is an isolate language (presumed Papuan) spoken by approximately 7,000 people residing on Rossel Island, an island found at the far end of the Louisiade Archipelago in Milne Bay Province, Papua New Guinea. The Yélî sound system, much like its baroque grammatical system (Levinson, 2020), is unlike any other in the region.

With only four primary places of articulation (bilabial, alveolar, post-alveolar, and velar) and no voicing contrasts, the phonological inventory is remarkably packed with acoustically similar segments. The core oral stop set includes both singleton (/p/, /t/, /ṭ/, and /k/) and doubly-articulated (/tp/, /ṭp/, /kp/) segments, with full nasal equivalents (/m/, /n/, /ṇ/, /ŋ/, /nm/, /ṇm/, /ŋm/), and with a substantial portion of them contrastively pre-nasalized or nasally released (/mp/, /nt/, /ṇṭ/, /ŋk/, /nmtp/, /ṇmṭp/, /ŋmkp/, /ṭṇ/, /kŋ/, /ṭpṇm/, /kpŋm/). A large number of this combinatorial set can further be contrastively labialized, palatalized on release, or both (e.g., /pʲ/, /pʷ/, /pʲʷ/; /tpʲ/; /ṇmḍbʲ/; see Levinson (2020) for details).\footnote{We use Levinson's (2020) under-dot notation (e.g., /ṭ/) to indicate the post-alveolar place of articulation; these stops are, articulatorily, somewhat variable in place, with at least some tokens produced fully sub-apically. In approximating cross-linguistic segment frequency below we use the corresponding retroflex for each stop segment (e.g., /ʈ/, /ʈp/, /ɳ/).} The consonantal inventory also includes a number of non-nasal continuants (/w/, /j/, /ɣ/, /l/, /βʲ/, /lʲ/, /lβʲ/). Vowels in Yélî Dnye may be oral or nasal, short or long. The 10 oral vowel qualities, which span four levels of vowel height, (/i/, /ɯ/, /u/, /e/, /o/, /ə/, /ɛ/, /ɔ/, /æ/, /ɑ/) can be produced as short and long vowels, with 7 of these able to appear as short and long nasal vowels as well /ĩ/, /ũ/, /ə̃/, /ɛ̃/, /ɔ̃/, /æ̃/, /ɑ̃/).

In total, Yélî Dnye uses 90 distinctive segments (not including an additional 3 rarely used consonants), far outstripping the phonemic inventory size of other documented Papuan languages (Foley, 1986; Levinson, 2020) and includes at least two contrasts that had not previously been documented elsewhere (labial-coronal double-articulations with dental vs.~post-alveolar coronal placement for both oral and nasal stops, Ladefoged \& Maddieson, 1996; Maddieson \& Levinson, n.d.).

Most words in Yélî Dnye are bisyllabic (\textasciitilde{}50\%), with monosyllabic words (\textasciitilde{}40\%) appearing most commonly after that, and with tri-and-above syllabic words appearing least frequently (\textasciitilde{}10\%; based on \textgreater{}5800 lexemes in the most recent Yélî Dnye dictionary at the time of writing; Levinson, 2020). The vast majority of syllables use a CV format. A small portion of the lexicon features words with a final closed syllable, limited to codas of -/m/, -/p/, or -/j/ (e.g., \enquote{ndap} (Spondylus shell) /ṇṭæp/). However, in spontaneous speech an epenthetic /ɯ/ is often inserted after word-final coda, yielding a CV.CV structure instead (e.g., \enquote{ndap} pronounced \enquote{ndapî} /'ṇṭæ.pɯ/). This process is used frequently with English loan words that have a coda (e.g., \enquote{ponî} (phone) /'pɔ.nɯ/). An even smaller portion of the lexicon features words starting with a vowel (e.g., \enquote{ala} (here) /æ.'læ/), but these are limited to /æ/-. Finally, the lexicon features a handful of single-vowel grammatical morphemes (see Levinson (2020) for details).

@@

In a basic NWR task, the participant listens to a production of a word-like form, such as /bilik/, and then repeats back what they heard without changing any phonological feature that is contrastive in the language. For instance, in English, a response of {[}bilig{]} or {[}pilik{]} would be scored as incorrect; a response {[}biːlik{]}, where the vowel is lengthened without change of quality would be scored as correct, because English does not have a general feature of contrastive vowel length. There is some variation in how past NWR studies have designed the presentation procedure and structure of items. For example, while items are often presented orally by the experimenter (Torrington Eaton, Newman, Ratner, \& Rowe, 2015), an increasing number of studies have turned instead to playing back pre-recorded stimuli in order to increase control in stimulus presentation (Brandeker \& Thordardottir, 2015). Additionally, while some studies have used 10-15 non-words, others have employed up to 46 unique items (Piazzalunga et al., 2019). Authors also often modulate structural complexity, typically measured in terms of item length (measured in number of syllables) and/or syllable structure (open as opposed to closed syllables, Gallon et al., 2007).

@@

We had considered boosting the interpretational value of this evidence by announcing our analysis plans prior to conducting them. However, we realized that even pre-registering an analysis would be equivocal because we would not have enough power to look at all relationships of interest, in many cases possibly not enough to detect any of the known effects, given their variability across studies. To illustrate this issue, we portray in Figure 1 studies in which children's NWR scores were gathered between 4 and 12 years of age, and reported separately for items that are relatively short (1-2 syllables) versus longer items (3-4 syllables). Notice that the effect of stimulus length is minuscule among Italian children (Piazzalunga et al., 2019), but considerable among Tsimane' children (Cristia et al., 2020), where a drop of 40 percentage points is observed at all ages. A similar large difference in NWR scores for short versus long items was observed among Arabic children (Jaber-Awida, 2018). Even the effect of age is unstable in this sample of studies. Whereas it is quite clear that children's NWR scores increase in the Italian data, age effects are less stable among Tsimane' children. Therefore, all analyses in the present study are descriptive and should be considered exploratory.

@@

A second list of candidate items was generated in 2019 by selecting complex consonants and systematically crossing them with all the vowels in the Yélî Dnye inventory to produce consonant-vowel monosyllabic forms. As before, items were automatically excluded if they appeared in the dictionary. Additionally, since perceiving vowel length in isolated monosyllables is challenging, any item that had a short/long lexical neighbor was excluded. Because there is still much to discover about the phonology and phonetics of Yélî Dnye (Levinson, 2020), it was also possible that we initially generated items with illegal, but currently undocumented constraints. Therefore, we made sure that the precise consonant-vowel sequence occurred in some real word in the dictionary (i.e., that there was a longer word included the monosyllable as a subsequence). These candidates were then presented to one informant, for a final check that they did not mean anything. Together with the 2018 selection, they were recorded, based on their orthographic forms, using a Shure SM10A XLR dynamic headband microphone and an Olympus WS-832 stereo audio recorder (using an XLR to mini-jack adapter) by the same informant, monitored by the second author for clear production of the phonological target. The complete recorded list was finally presented to two more informants, who were able to repeat all the items and who confirmed there were no real words present. Despite these checks, one monosyllable was ultimately frequently identified as a real word in the resulting data (intended \enquote{yî} /yɯ/; identified as \enquote{yi} /yi/, \emph{tree}). This item is removed from the analyses below.

@@

We ran children in batches, testing within a handful of hamlet clusters spread across the northeastern region of the island. Because space availability was limited in different ways from hamlet to hamlet, the places where elicitation happened varied across testing sites. We tested in four different sites, only making a single visit to each, conducting back-to-back testing of all eligible children present at the time of our visit in order to prevent the items from \enquote{spreading} between children through hearsay. In the first hamlet, we tested children in five different places, with some children being tested inside a house and others tested on the veranda. More information is available from the online supplementary materials. Whenever children living in the same household were tested, we tried to test children in age order, from oldest to youngest, to minimize intimidation for younger household members, and always using different elicitation sets.

@@

Additionally, we estimated the typological frequency of all phonological segments used in the target items using the PHOIBLE cross-linguistic phonological inventory database (Moran \& McCloy, 2019). For each phone in our task, we extracted the number and percentage of languages noted to have that phone in its inventory. While PHOIBLE is an unprecedentedly comprehensive database, with phonological inventory data for over 2000 languages at time of writing, it is of course still far from complete, so estimated frequencies should be taken with a pinch of salt. In both PHOIBLE and UPSID---another large phonemic inventory database---note that nearly half of the segment types are only attested in one language (Steven Moran, personal communication; see also \href{http://web.phonetik.uni-frankfurt.de/upsid_info.html}{web.phonetik.uni-frankfurt.de/upsid\_info.html}). It is therefore unsurprising that we did not find three of our segments (/lβʲ/, /ṭp/, and /tp/) in the database. Following the general pattern that many phones only appear once cross-linguistically, even in large databases, we treat these unattested phones as having a frequency of 1, with a (rounded) percentage of 0\%.

@@

We next checked whether the association between whole-item NWR scores and cross-linguistic phone frequence could actually be due to frequency of the sounds within the language: One can suppose that sounds that occur more frequently across languages are also more frequent within a language, and therefore may be easier for children to represent and repeat. We estimated frequency of the phones present in the stimuli in a corpus of child-centered recordings (Casillas et al., 2020) by counting the number of word types in which they occurred, and applied the natural logarithm.\footnote{We also carried out this analysis using token phone frequency, but this measure was not correlated with whole-item NWR scores, and therefore the fact that it did not explain away the predictive value of cross-linguistic phone frequency was less informative than the relationship discussed in the main text, with type frequencies.} Here, unattested sounds were not considered (i.e., they were declared NA so that they do not count for analyses). Phone frequencies estimated from this corpus were correlated with cross-linguistic phone frequencies (r(27)=0.50, p \textless{} 0.01). Moreover, averaging across items and participants, the Pearson correlation between scaled average corpus phone frequency and whole-item NWR scores was r = 0.46. Therefore, we fit another mixed logistic regression, this time declaring as fixed effects both scaled cross-linguistic and corpus frequencies (averaged across all attested phones within each stimulus item), in addition to age. As before, the model contained random slopes for both child ID and target. In this model, both cross-linguistic frequency (ß = 0.82, SE ß = 0.27, p \textless{} 0.01) and age (ß = 0.33, SE ß = 0.13, p = 0.01) were significant predictors of whole-item NWR scores, but corpus frequency (ß = 0.01, SE ß = 0.25, p = 0.98) was not.

@@

Finally, we looked into the frequency with which mispronunciations resulted in real words. In fact, two thirds of incorrect repetitions were recognizable as real words or phrases in Yélî Dnye or English: 63\%. This type of analysis is seldom reported---we could only find one comparison point: Castro-Caldas, Petersson, Reis, Stone-Elander, and Ingvar (1998) found that illiterate European Portuguese adults' NWR mispronunciations resulted in real words in 11.16\% of cases, whereas literate participants did so in only 1.71\% of cases. The percentage we observe here is much higher than reported in Castro and colleagues' study, but we do not know whether age, language, or even test structure explains this difference.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-balladares2016socio}{}%
Balladares, J., Marshall, C., \& Griffiths, Y. (2016). Socio-economic status affects sentence repetition, but not non-word repetition, in Chilean preschoolers. \emph{First Language}, \emph{36}(3), 338--351. \url{https://doi.org/10.1177/0142723715626067}

@@

\leavevmode\hypertarget{ref-castro1998illiterate}{}%
Castro-Caldas, A., Petersson, K. M., Reis, A., Stone-Elander, S., \& Ingvar, M. (1998). The illiterate brain. Learning to read and write during childhood influences the functional organization of the adult brain. \emph{Brain: A Journal of Neurology}, \emph{121}(6), 1053--1063. \url{https://doi.org/10.1093/brain/121.6.1053}

@@

\leavevmode\hypertarget{ref-farabolini2021nonword}{}%
Farabolini, G., Rinaldi, P., Caselli, C., \& Cristia, A. (2021). Non-word repetition in bilingual children: The role of language exposure, vocabulary scores and environmental factors. \emph{Speech Language and Hearing}.
